<body style="background-color:lightblue;">


---
title: Spatial Regression Assignment
output: html_document
fig_width: 6 
fig_height: 4 
---




```{r library, include=FALSE}
library(biscale)
library(cleangeo)
library(cowplot)
library(dplyr)
library(geosphere)
library(ggplot2)
library(maps)
library(maptools)
library(rgdal)
library(rgeos)
library(sf)
library(sp)
library(spatialreg)
library(spdep)
library(tidyr)
```

```{r data_import}
data <- read.csv('./Data/childpov18_southfull.csv', 
                 colClasses = c("character", "character", "character", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric", 
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric"))


```



```{r view_data}
View(data)
```



```{r fix_names}
names(data)[names(data)=="X2016.child.poverty"] <- "child.pov.2016"
```

Subset data for information from the state of Florida 

```{r call-in_florida}
fl_pov <- data %>% subset(State == "FL")

```

checking data for any missing data or errors

```{r summary_fl}
summary(fl_pov)
```



# First let's start with Ordinary Least Squares


```{r OLS}
equation <- child.pov.2016 ~ rural + urban + lnmanufacturing + lnag + 
  lnretail + lnhealthss + lnconstruction + lnlesshs + 
  lnunemployment + lnsinglemom + lnblack + lnhispanic + 
  lnuninsured + lnincome_ratio + lnteenbirth + lnunmarried
```

limit set for numbers greater than 5 decimal places in the output and summaries

```{r options}
options(scipen = 5)
```

```{r run_OLS_test}
ols <- lm(equation, data=fl_pov)
summary(ols)
```


#Creating a list of contiguity neighbors
Used to determine any spatial relationships in the residuals. 
In order to create a county polygon using data from Federal Information Processing Standards codes (FIPS). 
```{r fips codes}

fips <- county.fips
fips.codes <- separate(data = fips, col = polyname, into = c("state", "county"), sep = ",")
fl_fips <- subset(fips.codes, state=="florida", select=fips)

florida <- map(database = "county", regions = "florida", fill=T, plot=F)
fl_sp = map2SpatialPolygons(florida,fl_fips$fips,CRS("+proj=longlat"))

```


```{r require_maptools}
require(maptools)
sf::sf_use_s2(FALSE)
```

cleangeo like janitor, but for spatial data

```{r cleangeo}
file <- system.file("extdata", "example.shp", package = "cleangeo")
sp <- readShapePoly(file)
 
 cleaned <- clgeo_Clean(fl_sp)
 report.clean <- clgeo_CollectionReport(cleaned)
 clgeo_SummaryReport(report.clean)
```

```{r neighbor_data}
neighb.data <- poly2nb(cleaned, queen=T)
cont.neighb <- nb2listw(neighb.data,style="W", zero.policy = TRUE)
```




Now that we have created a spatial dataset and a list of neighbors, we can determine if there is any residual spatial dependence.
# Moran's Correlation and LaGrange Multiplier Tests

examines residuals of the OLS regression with a sptial relationship matrix


```{r morans_correlation}
lm.morantest(ols, cont.neighb)
```
Looking above, we can see a significant p-value of 0.007844, so we would reject the null hypothesis that says there is no spatial correlation in the residuals.

**This would suggest we should be using a spatial model.


#LaGrange Multiplier Test
```{r LaGrange_Test}
lm.LMtests(ols, cont.neighb, test="all")
```
The results show: 

LMerr       LMlag        RLMerr     RLMlag    SARMA

0.1008      0.2543      0.2167      0.7156     0.2435      




Closest model would be LMerr with a p-value of 0.1008, but is still high 



# Spatially lagged X Model

```{r SLX}
SLX.model <- spatialreg::lmSLX(equation, data=fl_pov, cont.neighb)
summary(SLX.model)
```
Note p-value of this model is significant, but no variables stick out here, except retail with a p-value of 0.079 


```{r SLX_summary}
summary(spatialreg::impacts(SLX.model, cont.neighb), zstats = TRUE)[["pzmat"]]
```

Not much to see above. 


#Spatial Lag Model
```{r spatial lag model}
sp.lag.model <- spatialreg::lagsarlm(equation, data=fl_pov, cont.neighb)
summary(sp.lag.model, Nagelkerke = TRUE)
```

```{r summary_spatial lag}
summary(spatialreg::impacts(sp.lag.model, listw = cont.neighb, R=100), zstats = TRUE)[["pzmat"]]
```
This model did have a significant p-value of 0.0426


#Spatial Error Model

```{r aptial_error_model}
sp.err.model <- spatialreg::errorsarlm(equation, data=fl_pov, cont.neighb)
summary(sp.err.model, Nagelkerke = TRUE)
```
Note significant p-value






```{r Hausman_test}
spatialreg::Hausman.test(sp.err.model)
```

The Hausman test resulted in a high p-value of 0.5553, which would fail to reject the null hypothesis that spatial error model is needed. 



#Spatial Durbin Error Model
The Spatial Durbin Error Model contains components of OLS, SLX, and Spatial Error models.

```{r Durbin_test}

sd.err <- spatialreg::errorsarlm(equation, fl_pov, cont.neighb, etype = "emixed")
sdm <- spatialreg::lagsarlm(equation, fl_pov, cont.neighb, type = "mixed")
```

Remembering that the Spatial Durbin Error Model is a local model which includes the errors and lag x values, we can view a summary of that model.


```{r summary_Durbin_model}
summary(sd.err, Nagelkerke = TRUE)
```
Model not significant with a p-value of 0.17444, but we do see some significance with variables: construction, income ratio, retail, and healthss

Examining the impacts matrix as we have before we can determine if the Spatial Durbin Error Model is the most appropriate model for our data or if we should restrict the model to a spatial error, SLX, or OLS model.

```{r summary_dem}

summary(spatialreg::impacts(sd.err, listw = cont.neighb, R = 100), zstats = TRUE)[["pzmat"]]
```

*** This model did however, produce significant results for variables retail and healthss, 


This will help us test the HO that we should restrict the model to a more simple model. The first step is to determine if we should restrict the model from a Spatial Durbin Error model to a spatial error mode.

```{r spatial_error_model}
LR.Sarlm(sd.err,sp.err.model)
```
For this model, we would fail to reject the null hypothesis with a p-value of 0.079





#Creating a list of K-neighbors

reate xy data from the polygons

```{r k-neighbors}
all.xy <-centroid(fl_sp)
colnames(all.xy) <- c("x","y")
```



We will examine k = 1, k = 3, and k = 5. Then we need to calculate the distance value so the model can create a radius to encompass the neighbors. Finally, we need to produce the list of neighbors within the neighborhood.



```{r neighbors}
#Create neighbors
all.dist.k1 <- knn2nb(knearneigh(all.xy, k=1, longlat = TRUE))
all.dist.k3 <- knn2nb(knearneigh(all.xy, k=3, longlat = TRUE))
all.dist.k5 <- knn2nb(knearneigh(all.xy, k=5, longlat = TRUE))
all.dist.k9 <- knn2nb(knearneigh(all.xy, k=9, longlat = TRUE))

#Determine max k distance value to neighbor
all.max.k1 <- max(unlist(nbdists(all.dist.k1, all.xy, longlat=TRUE)))
all.max.k3 <- max(unlist(nbdists(all.dist.k3, all.xy, longlat=TRUE)))
all.max.k5 <- max(unlist(nbdists(all.dist.k5, all.xy, longlat=TRUE)))
all.max.k9 <- max(unlist(nbdists(all.dist.k9, all.xy, longlat=TRUE)))

#Calculate neighbors based on distance
all.sp.dist.k1 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k1, longlat = TRUE)
all.sp.dist.k3 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k3, longlat = TRUE)
all.sp.dist.k5 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k5, longlat = TRUE)
all.sp.dist.k9 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k9, longlat = TRUE)

#Create neighbor list
all.dist.neighb.k1 <- nb2listw(all.sp.dist.k1,style="W", zero.policy = TRUE)
all.dist.neighb.k3 <- nb2listw(all.sp.dist.k3,style="W", zero.policy = TRUE)
all.dist.neighb.k5 <- nb2listw(all.sp.dist.k5,style="W", zero.policy = TRUE)
all.dist.neighb.k9 <- nb2listw(all.sp.dist.k9,style="W", zero.policy = TRUE)
```





#Distance Lab Model

calculating distance for k=1, k=3, and k=5

```{r distance_lag_model}
all.dist.lag.k1 <- spatialreg::lagsarlm(equation, data = fl_pov, listw = all.dist.neighb.k1)
all.dist.lag.k3 <- spatialreg::lagsarlm(equation, data = fl_pov, listw = all.dist.neighb.k3)
all.dist.lag.k5 <- spatialreg::lagsarlm(equation, data = fl_pov, listw = all.dist.neighb.k5)
all.dist.lag.k9 <- spatialreg::lagsarlm(equation, data = fl_pov, listw = all.dist.neighb.k9)
```


Summary for k=1

```{r k=1}
summary(all.dist.lag.k1, Nagelkerke = TRUE)
```
For k=1, there is a high p-value of 0.455.




Now summary for k=3

```{r k=3}
summary(all.dist.lag.k3, Nagelkerke = TRUE)
```
Another high p-value when looking at k=3





summary for k=5

```{r k=5}
summary(all.dist.lag.k5, Nagelkerke = TRUE)
```

Another high p-value when looking at k=5






Summary for k=9

```{r k=9}
summary(all.dist.lag.k9, Nagelkerke = TRUE)
```

For k=9, there is a significant p-value of 0.03622



#K-distance Error Model

Run for neighbors 1,3,5, and 9


```{r K-distance error model}

all.dist.err.k1 <- spatialreg::errorsarlm(equation, data = fl_pov, listw = all.dist.neighb.k1)
all.dist.err.k3 <- spatialreg::errorsarlm(equation, data = fl_pov, listw = all.dist.neighb.k3)
all.dist.err.k5 <- spatialreg::errorsarlm(equation, data = fl_pov, listw = all.dist.neighb.k5)
all.dist.err.k9 <- spatialreg::errorsarlm(equation, data = fl_pov, listw = all.dist.neighb.k9)
```









```{r summary_k-distance error model}
summary(all.dist.err.k1, Nagelkerke = TRUE)
```
This model with k=1 was significant with a p-value of 0.0169, the closest variable is construction.



Distance     Lag     Model    Distance   Err    Model






#Mapping the results for child poverty and Retail Work

```{r output_data}
dist.err.data <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output <- cbind.data.frame(fl_pov$FIPS,
                                    dist.err.data$fitted.values, 
                                    dist.err.data$residual, 
                                    fl_pov$child.pov.2016, 
                                    fl_pov$lnretail, 
                                    fl_pov$lnhealthss, 
                                    fl_pov$lnincome_ratio,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output) <- c("fips","fitted","resid","childpov",
                        "Retail","Healthss","Income Ratio")
```
                        
                        
```{r fortify}                        
                        
                        
fl_fortify <- fortify(fl_sp)

fl_poly <- merge(x = fl_fortify, y = dist.err.output, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data <- bi_class(fl_poly, x = childpov, y = Retail, 
                           dim = 3, style = "quantile")

legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Retail Work",
                    size = 6)
                    
                    
```                    
                    


```{r subset_us_mapppp}                 

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))
                              
```

```{r mapppp}                             
retail_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = bivariate_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "GrPink", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-88,-80), ylim=c(24,31)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Retail Work") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
retail_pov_map
```

```{r added legend}
legend_map <- ggdraw() +
  draw_plot(retail_pov_map, 0, 0, 1, 1) +
  draw_plot(legend, -0.000000444, 0.12, 0.6, 0.30)
legend_map
```


#Mapping the results for child poverty and Healthss

```{r output_data1}
dist.err.data1 <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output1 <- cbind.data.frame(fl_pov$FIPS,
                                    dist.err.data1$fitted.values, 
                                    dist.err.data1$residual, 
                                    fl_pov$child.pov.2016, 
                                    fl_pov$lnretail, 
                                    fl_pov$lnhealthss, 
                                    fl_pov$lnincome_ratio,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output1) <- c("fips","fitted","resid","childpov",
                        "Retail","Healthss","Income Ratio")
```
                        
                        
```{r fortify1}                        
                        
                        
fl_fortify1 <- fortify(fl_sp)

fl_poly <- merge(x = fl_fortify, y = dist.err.output1, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data1 <- bi_class(fl_poly, x = childpov, y = Healthss, 
                           dim = 3, style = "quantile")

legend1 <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Health SS",
                    size = 6)
                    
                    
```                    
                    


```{r subset_us_mapppp1}                 

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))
                              
```

```{r mapppp1}                             
healthss_pov_map1 <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = bivariate_data1, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "GrPink", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-88,-80), ylim=c(24,31)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Health SS") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
healthss_pov_map1
```


```{r added legend1}
legend_map1 <- ggdraw() +
  draw_plot(healthss_pov_map1, 0, 0, 1, 1) +
  draw_plot(legend1, -0.000000444, 0.12, 0.6, 0.30)
legend_map1
```
















#Mapping the results for child poverty and Income Ratio

```{r output_data2}
dist.err.data2 <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output2 <- cbind.data.frame(fl_pov$FIPS,
                                    dist.err.data2$fitted.values, 
                                    dist.err.data2$residual, 
                                    fl_pov$child.pov.2016, 
                                    fl_pov$lnretail, 
                                    fl_pov$lnhealthss, 
                                    fl_pov$lnincome_ratio,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output2) <- c("fips","fitted","resid","childpov",
                        "Retail","Healthss","Income_Ratio")
```
                        
                        
```{r fortify2}                        
                        
                        
fl_fortify2 <- fortify(fl_sp)

fl_poly2 <- merge(x = fl_fortify, y = dist.err.output2, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data2 <- bi_class(fl_poly2, x = childpov, y = Income_Ratio, 
                           dim = 3, style = "quantile")

legend2 <- bi_legend(pal = "DkViolet",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "Income Ratio",
                    size = 6)
                    
                    
```                    
                    


```{r subset_us_mapppp2}                 

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))
                              
```

```{r mapppp2}                             
Income_pov_map2 <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = bivariate_data2, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "DkViolet", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-88,-80), ylim=c(24,31)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Income Ratio") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
Income_pov_map2
```

```{r added legend2}
legend_map2 <- ggdraw() +
  draw_plot(Income_pov_map2, 0, 0, 1, 1) +
  draw_plot(legend2, -0.000000444, 0.12, 0.6, 0.30)
legend_map2
```





#Mapping the results for child poverty and Construction

```{r output_data3}
dist.err.data3 <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output3 <- cbind.data.frame(fl_pov$FIPS,
                                    dist.err.data3$fitted.values, 
                                    dist.err.data3$residual, 
                                    fl_pov$child.pov.2016, 
                                    fl_pov$lnretail, 
                                    fl_pov$lnhealthss, 
                                    fl_pov$lnincome_ratio,
                                    fl_pov$lnconstruction,
                                    stringsAsFactors = FALSE)

#Renaming columns
colnames(dist.err.output3) <- c("fips","fitted","resid","childpov",
                        "Retail","Healthss","Income_Ratio","construction")
```
                        
                        
```{r fortify3}                        
                        
                        
fl_fortify3 <- fortify(fl_sp)

fl_poly3 <- merge(x = fl_fortify3, y = dist.err.output3, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data3 <- bi_class(fl_poly3, x = childpov, y = construction, 
                           dim = 3, style = "quantile")

legend3 <- bi_legend(pal = "Brown",
                    dim = 3,
                    xlab = "Child Poverty",
                    ylab = "In Construction",
                    size = 6)
                    
                    
```                    
                    


```{r subset_us_mapppp3}                 

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))
                              
```

```{r mapppp3}                             
Construction_pov_map3 <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "white") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "gray", color = "white") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "white") +  
  geom_polygon(data = bivariate_data3, aes(x=long, y=lat, group=group, fill = bi_class), color = "grey50", show.legend = FALSE) + 
  bi_scale_fill(pal = "Brown", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-88,-80), ylim=c(24,31)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'deepskyblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Bivariate Map of Child Poverty and Construction Work") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
Construction_pov_map3
```

```{r added legend3}
legend_map3 <- ggdraw() +
  draw_plot(Construction_pov_map3, 0, 0, 1, 1) +
  draw_plot(legend3, -0.000000444, 0.12, 0.6, 0.30)
legend_map3
```






